\textls[-0]{\thetitle}

### 2023-03-06 canonicitymodel.tex
Working on canonicity model.

We follow the reducibility argument from~\cite{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} to construct the canonicity model.
Each of their proof is based on QIIT, Category with Family\cite{dybjer1995internal} and Generalized Algebraic Theory \cite{cartmell1986generalised} as the syntax. However, our proof will still work directly on the earlier introdcued syntax for readability.
We will first specify the formulation of the most basic canonicity theorem.

@@

\begin{align*}
  \denotesC{\goodCtx{\Gm}{i}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
  &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
  \denotesC{\goodType{\Gm}{T}{}} & \text{ is a function maps to Set s.t.} \\
  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a Set of closed terms } \goodTerm{\cdot}{\_}{\sub{T}{\gm}} \\ 
  &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
  \denotesC{\goodSub{\Gm}{\gm}{\Delta}} & \text{ is a function}: \denotesC{\goodCtx{\Gm}{i}} \to \denotesC{\goodCtx{\Delta}{i}} \text{ s.t. } \denotesC{\goodSub{\Gm}{\gm}{\Delta}}(\sigma) \coloneqq \SubstComp{\sigma}{\gm} \\
  &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
  \denotesC{\goodTerm{\Gm}{t}{T}} & \text{ is a dependent function}: \prod_{\gm \in \denotesC{\goodCtx{\Gm}{i}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)  \text{ s.t. } \denotesC{\goodTerm{\Gm}{t}{T}}(\gm) \coloneqq \sub{t}{\gm} \\
  \denotesC{\goodType{\Gm}{\top}{}}(\gm) &= \{t : t \equiv \unit \} \\
  \denotesC{\goodType{\Gm}{\bot}{}}(\gm) &= \emptyset \\
  \denotesC{\goodType{\Gm}{\cB}{}}(\gm) &= \{t : t \equiv \true \text{ or } t \equiv\false \} \\
  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm) &= \{t : t \equiv \eqrefl{\sub{a}{\gm}} \} \text{ when } \sub{a}{\gm} \equiv \sub{b}{\gm} \\ 
  &= \emptyset \text{ otherwise } \\ 
  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}} &= \\
  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}} &= \\
  \denotesC{\goodCtx{\cdot}{}} &= \{\SubstEmp\} \text{  the singleton set with identity substitution} \\ 
  \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\app{t}}{B}} &= \\
  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &=
\end{align*}

@@

Now that we construct the above model, we can map an arbitrary closed term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \sub{\cB}{\SubstEmp} = \{\true, \false\}$, Thus $t \equiv \true \text{ or } t \equiv\false$.

>>>>

Now we prove the consistency and canonicity using canoncity model. We follow the reducibility argument from~\cite{kaposi2019gluing, coquand2018canonicity,sterling2019algebraic} to construct the canonicity model.
Each of their proof is based on QIIT, Category with Family\cite{dybjer1995internal} and Generalized Algebraic Theory \cite{cartmell1986generalised} as the syntax. However, our proof will still work directly on the earlier introdcued syntax for readability.
We will first specify the formulation of the most basic canonicity theorem.

@@

\begin{align*}
  \denotesC{\goodCtx{\Gm}{i}} & \text{ is a Set of closed substitution} \goodSub{\cdot}{\_}{\Gm} \\
  &\text{and we define } \goodCtxS{\Gm}{i} :\iff \denotesC{\goodCtx{\Gm}{i}} \text{ is defined} \\
  \denotesC{\goodType{\Gm}{T}{}} & \text{ is a function maps to Set s.t.} \\
  & \text{for arbitrary } \gm \in \denotesC{\goodCtx{\Gm}{}}, \denotesC{\goodType{\Gm}{T}{}}(\gm) \text{ is a Set of closed terms } \goodTerm{\cdot}{\_}{\sub{T}{\gm}} \\ 
  &\text{and we define } \goodTypeS{\Gm}{T}{i} : \iff \denotesC{\goodType{\Gm}{T}{i}} \text{ is defined } \\ 
  \denotesC{\goodSub{\Gm}{\gm}{\Delta}} & \text{ is a function}: \denotesC{\goodCtx{\Gm}{i}} \to \denotesC{\goodCtx{\Delta}{i}} \text{ s.t. } \denotesC{\goodSub{\Gm}{\gm}{\Delta}}(\sigma) \coloneqq \SubstComp{\sigma}{\gm} \\
  &\text{and we define } \goodSubC{\Gm}{\gm}{\Delta} :\iff \denotesC{\goodSub{\Gm}{\gm}{\Delta}} \text{ is defined} \\
  \denotesC{\goodTerm{\Gm}{t}{T}} & \text{ is a dependent function}: \prod_{\gm \in \denotesC{\goodCtx{\Gm}{i}}}\denotesC{\goodType{\Gm}{T}{}}(\gm)  \text{ s.t. } \denotesC{\goodTerm{\Gm}{t}{T}}(\gm) \coloneqq \sub{t}{\gm} \\
  \denotesC{\goodType{\Gm}{\top}{}}(\gm) &= \{t : t \equiv \unit \} \\
  \denotesC{\goodType{\Gm}{\bot}{}}(\gm) &= \emptyset \\
  \denotesC{\goodType{\Gm}{\cB}{}}(\gm) &= \{t : t \equiv \true \text{ or } t \equiv\false \} \\
  \denotesC{\goodType{\Gm}{\TyId{a}{b}}{}}(\gm) &= \{t : t \equiv \eqrefl{\sub{a}{\gm}} \} \text{ when } \sub{a}{\gm} \equiv \sub{b}{\gm} \\ 
  &= \emptyset \text{ otherwise } \\ 
  \denotesC{\goodType{\Gm}{\TyPi{A}{B}}{}} &= \\
  \denotesC{\goodType{\Gm}{\TySigma{A}{B}}{}} &= \\
  \denotesC{\goodCtx{\cdot}{}} &= \{\SubstEmp\} \text{  the singleton set with identity substitution} \\ 
  \denotesC{\goodCtx{\Gm, T}{}} &= \{ (\gamma , t) : \gamma \in \denotesC{\goodCtx{\Gm}{}}, t \in \denotesC{\goodType{\cdot}{\sub{T}{\gm}}{}}  \} \\
  \denotesC{\goodTerm{\Gm}{\unit}{\top}} &= \\
  \denotesC{\goodTerm{\Gm}{\true}{\cB}} &= \\
  \denotesC{\goodTerm{\Gm}{\false}{\cB}} &= \\
  \denotesC{\goodTerm{\Gm}{\ifb{c}{a}{b}}{T}} &= \\
  \denotesC{\goodTerm{\Gm}{\eqrefl{t}}{\TyId{t}{t}}} &= \\
  \denotesC{\goodTerm{\Gm}{\lam{t}}{\TyPi{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\app{t}}{B}} &= \\
  \denotesC{\goodTerm{\Gm}{\pair{a}{b}}{\TySigma{A}{B}}} &= \\
  \denotesC{\goodTerm{\Gm}{\fst{t}}{T}} &= \\
  \denotesC{\goodTerm{\Gm}{\snd{t}}{T}} &= \\
\end{align*}

@@

The first observation is that, with this model, we know our system is consistent---we cannot derive $\goodTerm{\cdot}{t}{\bot}$. Otherwise, we have $\denotesC{\goodTerm{\cdot}{t}{\bot}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \denotesS{\bot}(\star) = \emptyset$, that is an element in an empty set, a contradiction.

\begin{theorem}[Consistency]
  We cannot derive a closed term of bottom type, i.e. ,$\cdot \not \vdash t : \bot$ for arbitrary $t$.
\end{theorem}

What's more, with the above model, we can map an arbitrary closed term $\goodTerm{\cdot}{t}{\cB}$ to get the result $\denotesC{\goodTerm{\cdot}{t}{\cB}}(\SubstEmp) = \sub{t}{\SubstEmp} \in \sub{\cB}{\SubstEmp} = \{\true, \false\}$, Thus $t \equiv \true \text{ or } t \equiv\false$, arriving at the basic canonicity result.
